Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-09-05 paper/paper.tex
minor edits, color on single plots, 350 second intervals on paper/plot Tested everything on quipu.
plotocide Tested everything on quipu.

\begin{figure}
  \includegraphics[width=\columnwidth]{stability-analysis}
  \caption{Infinite slab stability analysis yields characteristic half
    wavelengths at which system becomes unstable.}
    \label{fig:stability-analysis}
\end{figure}

@@

\begin{figure}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-8_00-8_00-95_00-15_00-exact/plots/box-plot_D}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-10_00-17_00-94_00-15_00-exact/plots/box-plot_D}
  \caption{Box plots of Mannik shapes for sizes at the lower stability
    threshhold.  The vertical axis shows stacked and changing over
    time the total number of proteins that are of four different
    compound stages and in the different sections of the cell.  Both
    shapes have longer axis of roughly 2.2$\mu m$.}
  \label{box-mannik}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-18_50-18_50-95_00-15_00-full_array/plots/box-plot_D}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-18_50-18_50-95_00-15_00-exact/plots/box-plot_D}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-18_60-28_60-94_00-15_00-full_array/plots/box-plot_D}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-18_60-28_60-94_00-15_00-exact/plots/box-plot_D}
  \caption{A comparison of our box plots that show a disparity between
    the behavoir of the stochastic and deterministic simulations in
    our Mannik cells.  The deterministic show a regular, robust
    oscillatory behavoir between the two ends of the cell. The
    stochastic simulation do show a general back and forth tendency
    between two ends of the cell, but it is not robust and tends to
    break down.}
  \label{box-mannik}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-stad/0_25-5_50-1_00-0_00-15_00-full_array/plots/plot-time-averaged-arrow-NflD-500-2000}
  \includegraphics[width=\columnwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/plot-time-averaged-arrow-NflD-500-2000}
  \caption{Arrows depicting global maxima in space and local maxima in
    time overlayed on color plots of time averaged total MinD
    density. A density threshold is chosen and only maxima above this
    threshold are shown. The plots serve to illistrate what is
    observed in our movies. \fixme{how much of time of data is
      shown?}.  Although there is variation in the specific location
    of maxima on each edge of the cell, the behavoir for both of these
    regular shapes is robust and exhibits a back and forth oscillatory
    manner.  The wildtype pill shape behavoir matches well with
    experiment.  Regular pancake shapes have not been observed
    experimentally, but our stochastic simulations would suggest that
    they would exhibit regular robust oscillations similar to the pill
    shapes.  This is in contrast to the behavoir of Mannik cells seen
    in Figure~\ref{randst-plot-ave}}
  \label{arrow-pill-stadium}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-stad/0_25-5_50-1_00-0_00-15_00-full_array/plots/NflD-time-map}
  \caption{Time map plots of our stad shape}
  \label{time-maps-stad}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-18_50-18_50-95_00-15_00-full_array/plots/NflD-time-map}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-18_60-28_60-94_00-15_00-full_array/plots/NflD-time-map}
  \includegraphics[width=\columnwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/NflD-time-map}
  \caption{Time map plots of our randst shape}
  \label{time-maps-randst}
\end{figure}

@@

%% \section{Triangular Shapes}

%% \begin{figure}
%% %  \includegraphics[width=4.2cm]{../data/shape-triangle/plots/box-plot_D--triangle-25-233-233-233-1500}
%% %  \includegraphics[width=4.2cm]{../data/shape-triangle/plots/box-plot_D--triangle-25-266-266-266-1500}
%%   \includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-233-233-233-1500}
%%   \includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-266-266-266-1500}
%%   %\includegraphics[width=4.2cm]{../data/shape-triangle/plots/box-plot_D--triangle-25-800-800-800-1500}
%%   \caption{Two box plots of triangle shapes.  The two smaller show
%%     sizes at the lower stability threshhold, one just below, one just
%%     above.  The larger triangle is the largest simulated. The vertical
%%     axis shows stacked and changing over time the total number of
%%     proteins that are of four different compound stages and in the
%%     different sections of the cell.}
%%   \label{box-triangle}
%% \end{figure}

%% Along with our other shapes we have chosen to simulate both
%% equilateral and iscoscolese triangular flattened shapes, since these
%% offer in a way the simplest shapes that are geometrically more
%% complicated than our pills.

%% Both triangle types show a similar oscillation pattern.  The
%% iscosolese exhibit regular oscillations characterized by one distinct
%% pole in the more accute angle of the three, and another 'pole' along
%% the opposite wall, peaking in the two more obtuse angles at the same
%% time.  The equilateral triangles settle very quickly into the same
%% pattern, where the choice of which is one pole corner is dependent
%% upon initial conditions.  As with the Mannik simulated shapes, there
%% is a critical small size below which the oscillations dampen out and
%% will not achieve a regular pattern.  Figure~\ref{box-triangle} shows
%% two plots of equilateral triangles.  The two smaller show the
%% largest that showed damped out behavoir and the smallest that
%% developed regular oscillations.  The larger shows the largest traingle
%% that we simulated, showing that at even very large sizes the behavoir
%% is consistent and stable.  Above the small size threshold, all
%% simulated triangles show regular patterns for as long as simulated
%% (roughly 800 seconds).

%% \section{Our '\textgreater' Shaped Cells}
%% \begin{figure}
%%   \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-525-525-9900-1500}
%%   \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-550-550-9900-1500}
%%   \label{fig:Vbox}
%% \end{figure}
%% \begin{figure}
%%   \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-600-600-9900-1500}
%%   \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-1400-1400-9900-1500}
%%   \label{fig:Vbox2}
%% \end{figure}

%% We have also simulated a flattened '\textgreater' shape that shows an
%% interesting size dependent pattern (seen in
%% Figures~\ref{fig:Vbox}~and~\ref{fig:Vbox2}).  When the shape is small enough
%% that the distance from the apex tip to the opposite curve is less than
%% our stability limit, a simple, two pole oscillation pattern develops
%% and is sustained (where the two poles are on opposite leftward corners
%% of the cell).  When the cell is increased in size so that this
%% distance is approximately equal to our stability value, the cell
%% develops a double oscillation in time, where the maxima oscillates
%% from the lower corner to the rightward apex, then \emph{back to the
%%   lower corner}, then to the apex again, then to the upper corner,
%% then back to the apex, then to the upper corner, then to the apex
%% again, then back to the lower corner, and continues this pattern
%% indefinately.  Figure~\ref{fig:Vbox} shows the smaller size, for which the
%% protein oscillates simply, and an intermediate size, which shows a
%% pattern for which one out of three spikes in the regular oscillation
%% is less significant.  This second pattern seems to be a precursor to
%% the double oscillation that seen in Figure~\ref{fig:Vbox2}.  The second
%% plot shown in Figure~\ref{fig:Vbox2} shows a slightly larger cell for
%% which the distance measured is larger than our stability limit
%% distance.  This simulation shows a random pattern.

%% A detailed analysis of a time series of protein contour plots shows
%% that the importance of spatial dimensions is based on the ability of
%% protein in different stages to diffuse neccesary distances before
%% undergowing certain types of reactions.  The double oscillation
%% pattern shows that every time the membrane bound MinD maximizes in the
%% upper corner as opposed to the rightward apex, it does so because the
%% MinE has not had enough time to diffuse to this upper corner, leaving
%% a MinE free zone.  This particular pattern is specific to this shape
%% and size, but in general its the restrictions set in place by the
%% rates at which proteins diffuse and the distances they cover before
%% reacting that dictate their oscillatory bahavoir.

%% Simulations of this shape in which the two planar dimensions are large
%% compared to those shown(roughly $9\micron$ across) show no stable
%% regular periodic pattern. Animations show that the MinD acts as a sort
%% of traveling cluster that is constantly being eaten away by MinE
%% behind it.  When this cluster enters a region of high curvature that
%% ends in a dead end corner it performs the polar maxima behavoir seen
%% in the other simulations, regrouping out away from the corner where it
%% travels in the opposite direction while being ``chased'' away by the
%% MinE that itself is currently maximizing in the corner.  The maxima
%% occur on non-polar walls often, and will often split into two roaming
%% clusters.

Figure \ref{total-oscillation-triangle-plot} shows the equivalent
motion and transference between compound states for the triangular
states as Figure \ref{total-oscillation-plot} does for the pill shaped
cells.  The equilateral triangular shows that in all three sections
there is a spike in wall attached MinD-ATP directly followed by a
spike in wall attached MinD-MinE-ATP.  The pill had shown a similar
pattern except that the mid section exhibited no such pronounced
spike.  Interestingly, the right and mid sections show spikes at
similar times when compared to the left section spikes, which seem to
oscillate in a manner shifted pi off from the other two.  This
behavoir shows itself in the animations, which show a relaxation into
a stable oscillation from a maxima in one corner to a double maxima in
the other two corners, and back again.  The left section heights are
roughly twice that of the right section heights.  The simulations are
started with higher concentrations of proteins in primarily the right
section, with some overlap into the middle section, so that perhpas
this oscillation pattern seems reasonable.  However, it's interesting
to note that the oscillations, started with assymmetric density, do
not naturally rotate around the corners of the triangle in a circular
pattern.

It seems as well that there is a desired oscillation maxima for the
particular shape and size of the cell, as can be seen by the
relaxation, over the first four periods, to a regular maximal height
of the right and left sections' total protein at the height of each spike.

\fixme{Note to make sure I'm interpreting the iscosolese triangle data
  correctly: the triangle sides are 5,5,3, and the higher density
  starts in the corner that is opposite the short side.  This corner
  is covered by the right section.  I've verified this by looking at
  the protein\_microscopy and drawing a little picture, which is on my
  desk...if you dare look for something on my desk!!!}  The iscosolese
triangle starts its simulation with a higher concentration of protein
in the section that is opposite the shorter side.  It quickly relaxes
into a stable oscillation pattern.  This pattern shows that right
corner section regularly has higher spikes in total protein than the
other sections, and that its spikes appear one pi removed from the
spikes for the other two sections.  This a similar behavoir to that
which is seen in the equilateral triangle, differening in that it
converges more quickly towards this behavoir and that it shows a
greater exageration in the height of protein buildup in the right most
section.

\fixme{Note on interpretation of 3-4-5 triangle: the higher density
  starts to the right of a vertical line that is parrallel to the
  second longest side in the triangle (the longer of the two
  non-hypotonuse sides).  This means the higher concentration covers
  the right angle and also the smaller angle in the triangle} The
3-4-5 triangle exhibits behavoir that I need more data to properly
observe.  I does not converge as quickly as that for the iscosolese or
equilateral triangles.

It seems that there is an approximate critical cell size for which the
MinD will act in these ``wandering cluster'' formations.  The cell is
large enough so that there is a loss of MinD correlation between the
opposite sides (the protein in the right corner at any time is too far
away from the left to immediately affect the protein there).

\fixme{running larger and smaller.  Guessing the larger will do pole
  and pole, so talk about in terms of you have huge, pole and pole,
  then once get small enough double oscillations develop.  Why do they
  do that?  And hen you can do image plot for the one with consistent
  double oscillations.  Can right now explain reason for double
  oscillations.  Look at todo.txt.  The slightly varied guassian shape,
  so a little assymetric, doesn't change double oscillation pattern.}

Contour plot frames from the animations show the second maxima in the
lower left corner (immediately following the one in the upper middle
corner) will reside for a bit longer than the left corner maxima
preceding it.  These longer residing maxima will then be followed by a
switch, to the right corner, with a short, less intensive maxima
traveling through the center.  The MinE shows the same pattern. The
MinE that was just concentrated on the left corner during the first
maxima, after releasing the MinD from the membrane there, diffuses up
and attaches to the MinD on the upper corner membrane and begins to
release it.  During this process the MinE starts at the edges and
moves in towards the center of the maxima.  The center of the cup is a
little further away so it attaches to these outer ridges first and
then moves in.  After it eats at the corners and starts moving in the
MinD eventually travels back to set up a maxima in the left corner.
Watching the cytoplasmic MinE, something very interesting: It as well
maximizes twice in the left corner, but during the MinD's first maxima
in the left corner, the MinE has a maxima in the upper corner, and
during the MinD's second left maxima, the MinE travels to the right
corner and has a maxima there.  At the same time (during the second
MinD left corner maxima), there is a large amount of cytoplasmic
MinD:ATP on the right side.  The MinE does this becuase during the
first MinD left corner maxima, the MinE has just come off a maxima in
the right corner.  So it diffuses up to the upper corner and sets up
there.  During the second MinD left corner maxima, the MinE has just
come off the upper corner maxima.  It will diffuse into the right
corner and left corner.  In the left corner it interacts with the MinD
bound to the membrane there.  In the right corner it sits and
eventually diffuses away, but at a somewhat slow pace.  \fixme{An idea
  is that once the MinD travels back from the left corner after its
  second maxima there, it is not able tot set up a maxima in the upper
  corner as easily, since there is left over MinE there, which has
  just traveled from the right corner.  The problem with this is that
  the MinE doesn't seem to be maximized in the upper corner when the
  MinD is traveling back up from the left.}  The other idea is simply
that because the minE is traveling all the way from the right corner,
the longer time it takes to get there simply means that it all
concentrates on the left, leaving the upp and right side completely
open for the build up of MinD.  \fixme{Still dont quite get it but
  take a look at the other simulations and come back.} Actually, you
can see this effect very slightly watching the 98 shape, between the
center cap and the right corner.

Simulations of this shape that has the same height and dimensions .6
times the size (cells roughly $1.6\mu m$ from top to bottom) do not
exhibit the same pattern.  In these the MinD maxima oscillate quickly
back and forth between the two lower corner poles.

Simulations in which the two planar dimensions are 2 and then 3 times
the size (cells roughly $6\micron$ and $9\micron$, respectively, from top
to bottom) show no stable regular periodic pattern \fixme{although
  they haven't really run long enough to exhibit much.  They certainly
  won't show the double oscillation pattern, however.}  The animations
of $6\micron$ cells show that the MinD acts as a sort of traveling
cluster that is constantly being eaten away be MinE behind it.  When
this cluster enters a region of high curvature that ends in a dead end
corner it performs the polar maxima behavoir seen in the other
simulations, regrouping out away from the corner where it travels in
the opposite direction while being ``chased'' away by the MinE that
has itself maximized in the corner.  If it has just maximized in the
top corner, it regroups traveling downward and then naturally splits
into two groups that travel out and form maxima in the two opposite
bottom corners.  After maximizing in these two corners the two groups
recombine again in the center and travel in some direction together as
one cluster.  Simulations of the $9\micron$ cells show a similar roaming
cluster.  Although here, the maxima occur on non-polar walls often,
and will often split into two roaming clusters that interact less
quickly and are less correlated than those seen in the $6\micron$ cell.

It seems that there is an approximate critical cell size for which the
MinD will act in these ``wandering cluster'' formations.  The cell is
large enough so that there is a loss of MinD correlation between the
opposite sides (the protein in the right corner at any time is too far
away from the left to immediately affect the protein there).

>>>>

@@

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-18_50-18_50-95_00-15_00-full_array/plots/box-plot_D}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-18_50-18_50-95_00-15_00-exact/plots/box-plot_D}
  \caption{A comparison of our box plots that show a disparity between
    the behavoir of the stochastic and deterministic simulations in
    our Mannik cells.  The deterministic show a regular, robust
    oscillatory behavoir between the two ends of the cell. The
    stochastic simulation do show a general back and forth tendency
    between two ends of the cell, but it is not robust and tends to
    break down.}
  \label{box-mannik}
\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-18_60-28_60-94_00-15_00-full_array/plots/box-plot_D}
  \includegraphics[width=\columnwidth]{../data/shape-randst/0_25-18_60-28_60-94_00-15_00-exact/plots/box-plot_D}
  \caption{A comparison of our box plots that show a disparity between
    the behavoir of the stochastic and deterministic simulations in
    our Mannik cells.  The deterministic show a regular, robust
    oscillatory behavoir between the two ends of the cell. The
    stochastic simulation do show a general back and forth tendency
    between two ends of the cell, but it is not robust and tends to
    break down.}
  \label{box-mannik}
\end{figure}

@@

@@
